### 2022-09-27 chapters/optimization/python-dir/python.tex
remove : extra figure

\subsubsection{Python \& Multiprocessing}
The purpose of this part is to study the impact of concurrency on the energy consumption for Python applications.
To do so, we run a simple code snippet that computes the sum of the first $N$ numbers using the standard concurrency libraries \texttt{multiprocessing} and \texttt{multihtreading}. Using Desktop  machine with Intel(R) Core(TM) i7-4800MQ CPU @ 2.70GHz with four cores and four hyperthreads, we run the following strategies.
%  N ==5e5
\begin{itemize}
    \item Sequentiel: and here we just compute the sum of the first $N$ numbers without any concurency
    \item Multithreading: and here we use the \texttt{ ThreadPoolExecutor} library to compute the sum of the first $N$ numbers, We devide $N$ by the number of threads and each thread will compute the sum of the numbers in its range. For our case we used 4 threads.
    \item multiprocessing : and in this case, we use the \texttt{ProcessPoolExecutor} library to calculate the sum of the first $N$ numbers. We divide $N$ by the number of processes, and each process sums the numbers in its range. In this situation, we employed 2,4,8,16 processes, correspondingly.
\end{itemize}

Figure~\ref{fig:python_multiprocessing_strategies} shows the results of the experiment. As one can notice
\begin{figure}[!hbt]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_multiprocessing}
    \caption{Energy behaviour resulting from diffrent python concurrency strategies (number of workers) }
    \label{fig:python_multiprocessing_strategies}
\end{figure}

Additionally, \Cref{fig:python_multiprocessing} reports on the energy consumption of Python applications when introducing parallelism.
Python applications are single-threaded systems constrained by the \emph{global lock system} (GLS).
However, due to the increase of cores/threads per CPU, many Python libraries started to take advantage of this hardware feature by allowing concurrent execution of Python processes.
For example, the multiprocessing library\footnote{\url{https://docs.python.org/3/library/multiprocessing.html}} spawns sub-processes to increase the degree of concurrency of Python applications.
As one can see in \Cref{fig:python_multiprocessing}, the energy consumption is correlated with the number of threads until reaching the limit of physical cores, when concurrent processes start to compete for the CPU.

\begin{figure}[hbt]
    \centering
    \includegraphics[width=\linewidth]{imgs/multiprocessing_energyvstime}
    \caption{Energy consumption of Python multiprocessing depending on the number of exploited threads.}
    \label{fig:python_multiprocessing}
\end{figure}

>>>>

\subsubsection{Python \& Multiprocessing}
The purpose of this part is to study the impact of concurrency on the energy consumption for Python applications.
To do so, we run a simple code snippet that computes the sum of the first $N$ numbers using the standard concurrency libraries \texttt{multiprocessing~\footnote{\url{https://docs.python.org/3/library/multiprocessing.html}}} and \texttt{multihtreading~\footnote{\url{https://docs.python.org/3/library/threading.html}}}. Using Desktop  machine with Intel(R) Core(TM) i7-4800MQ CPU @ 2.70GHz with four cores and four hyperthreads, we run the following strategies.
%  N ==5e5
\begin{itemize}
    \item Sequentiel: and here we just compute the sum of the first $N$ numbers without any concurency
    \item Multithreading: and here we use the \texttt{ ThreadPoolExecutor} library to compute the sum of the first $N$ numbers, We devide $N$ by the number of threads and each thread will compute the sum of the numbers in its range. For our case we used 4 threads.
    \item multiprocessing : and in this case, we use the \texttt{ProcessPoolExecutor} library to calculate the sum of the first $N$ numbers. We divide $N$ by the number of processes, and each process sums the numbers in its range. In this situation, we employed 2,4,8,16 processes, correspondingly.
\end{itemize}

\begin{figure}[!hbt]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_multiprocessing}
    \caption{Energy behaviour resulting from diffrent python concurrency strategies (number of workers) }
    \label{fig:python_multiprocessing_strategies}
\end{figure}

Figure~\ref{fig:python_multiprocessing_strategies} shows the results of the experiment. As one can notice

performance
- sequentiel took less time then mulit threading
- 4 process was the optimal one
-- for cpu power the multi threading reached 13 watts whule the 4 process multiprocessing was 42 watts
-- for global energy  the sequentiel one took the most, then the multithreading after that the muli processing , while keeping the 4 process always the optimial one

Additionally, \Cref{fig:python_multiprocessing} reports on the energy consumption of Python applications.
Python applications are single-threaded systems constrained by the \emph{global lock system} (GLS).
However, due to the increase of cores/threads per CPU, many Python libraries started to take advantage of this hardware feature by allowing concurrent execution of Python processes.
For example, the multiprocessing library\footnote{\url{https://docs.python.org/3/library/multiprocessing.html}} spawns sub-processes to increase the degree of concurrency of Python applications.
As one can see in \Cref{fig:python_multiprocessing}, the energy consumption is correlated with the number of threads until reaching the limit of physical cores, when concurrent processes start to compete for the CPU.
